Peer Assessment 1 by Allan Chen
===============================
          
First, we are going to load the data into R.

```{r load data}
ds <- read.csv("activity.csv")
```
      
  
## Part I

To make the data more accessible for our analysis, we are going to introduce a new column to the data, called "day".

```{r}
day <- rep(1:61, each = 288)
pds <- cbind(day = day, ds)
```

So now, the data looks something like this:

```{r}
head(pds)
tail(pds)
```

Now that the data is much easier to subset, we can calculate the total number of steps taken per day, ignoring all of the missing values. 

```{r}
totalsteps.list <- c()
for(i in 1:61) {
        pdsi <- subset(pds, day == i)
        totalsteps <- sum(pdsi$steps, na.rm = TRUE)
        totalsteps.list <- c(totalsteps.list, totalsteps)
}
```

Using the base plotting system, we can make a **histogram** of the total number of steps taken each day. 
Note that based on how we processed the data, Day 1 is October 1st, 2012 and Day 61 is November 30, 2012.

```{r}
barplot(totalsteps.list, names.arg=1:61, main = "Total Numer of Steps Per Day", ylab = "Total Number of Steps", xlab = "Day")
```

We can also calculate the **mean** and **median total number of steps taken per day**. 

```{r}
mean <- mean(totalsteps.list)
median <- median(totalsteps.list)
```

The **mean total number of steps taken per day** is `r mean`.    
The **median total number of steps taken per day** is `r median`.             

## Part II

To make a **time-series plot** of the average number of steps taken for a given interval, we are going to use the loop function. 

For each iteration, we are going to subset the data based on the "interval" column. Then, with each subset, we are going to calculate the average number of steps taken and add the value to the "averagesteps.list" vector. 

```{r}
averagesteps.list <- c()
for(i in ds$interval[1:288]) {
        ds.int <- subset(ds, interval == i)
        averagesteps <- mean(ds.int$steps, na.rm = TRUE)
        averagesteps.list <- c(averagesteps.list, averagesteps)
}
```

Then, we are going to construct a new data frame with two columns: "interval" and "averagesteps"

```{r}
new.ds <- data.frame(interval = ds$interval[1:288], averagesteps = averagesteps.list)
```

The resulting data frame looks something like this:
```{r}
head(new.ds)
tail(new.ds)
```

Now, we can visualize the result by plotting it using the base plotting system.

```{r}
plot(averagesteps ~ interval, new.ds, type = "l", main = "Time-Series Plot of the Average Number of Steps Taken", xlab = "Interval", ylab = "Average Steps")
```

Furthermore, we can find which 5-minute interval contains the **maximum number of steps**.

```{r}
max <- new.ds$interval[which.max(new.ds[,2])]
```
  
The **maximum number of steps** is in the 5-minute interval: `r max`. 

## Part III

### Question 1 

We can calculate the **total number of missing values** by:

```{r, results ='hide'}
sum <- sum(is.na(ds[,1]))
```

The result is: `r sum`. 

### Question 2

The plan is that we are going to **replace all of the missing values** with the *mean total number of steps taken per interval*. The steps are:

1. Calculate the mean total number of steps taken per day.

2.Divide the mean total number of steps taken per day by the number of intervals per day. 

3.Then, we are going to identify the missing values in the data and replace them with the calculated value in *2*. 

4.We have now created a data with all of the missing values filled in. 

### Question 3 

Previously, we calculated the mean total number of steps taken per day to be: 

```{r}
mean(totalsteps.list)
```

There are 288 intervals per day, so the *mean total number of steps taken per interval* is: 

```{r}
mean(totalsteps.list) / 288
```

```{r, echo = FALSE, results = 'hide'}
replace.with <- mean(totalsteps.list) / 288
```
We can **replace all of the missing values** with our choice of value `r replace.with` by: 

```{r}
ds[is.na(ds)] <- 32.48
```

Now, our data looks something like this: 

```{r}
head(ds)
tail(ds)
``` 

### Question 4

Before we start, we are going to introduce a column, called "day".

```{r}
day <- rep(1:61, each = 288)
pds <- cbind(day = day, ds)
```

Exactly like what we did in ***Part I***, we can calculate the total number of steps taken per day by:

```{r}
totalsteps.list <- c()
for(i in 1:61) {
        pdsi <- subset(pds, day == i)
        totalsteps <- sum(pdsi$steps, na.rm = TRUE)
        totalsteps.list <- c(totalsteps.list, totalsteps)
}
```

Using the base plotting system, we can make a **histogram** of the total number of steps taken each day. 
Note that based on how we processed the data, Day 1 is October 1st, 2012 and Day 61 is November 30, 2012.

```{r}
barplot(totalsteps.list, names.arg=1:61, main = "Total Numer of Steps Per Day", ylab = "Total Number of Steps", xlab = "Day")
```

Finally, we calculate the **mean** and **median total number of steps taken per day**. 

```{r}
mean(totalsteps.list)
median(totalsteps.list)
```

The impact of imputing missing data instead of ignoring it is that the mean total number of steps taken per day is now slightly larger (*9354.23* versus *10581*).

## Part IV

For this part, we are going to use the data that we have constructed in ***Part III***.

In order to make a **time-series plot for weekdays versus weekends**, we need to first convert the "date" column to the class "Date". Then, we are going to add a column, called "wdwe", which is a factor variable embeeded inside an *ifelse function*. Lastly, we are going to insert the "wdwe" column into the data. 

```{r}
ds$date <- as.Date(ds$date, format = "%Y-%m-%d")
ds$wdwe<- as.factor(ifelse(weekdays(ds$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))
```

Now, the data looks something like this:
```{r}
head(ds)
tail(ds)
```

Then, we are going to calculate the average number of steps taken per 5-minute interval, averaged over all weekday days. 

```{r}
averagesteps.list1 <- c()
for(i in ds$interval[1:288]) {
        ds.int1<- subset(ds, interval == i & wdwe == "Weekday")
        averagesteps1 <- mean(ds.int1$steps)
        averagesteps.list1 <- c(averagesteps.list1, averagesteps1) 
}
```

We are going to do the same for all weekend days. 

```{r}
averagesteps.list2 <- c()
for(i in ds$interval[1:288]) {
        ds.int2<- subset(ds, interval == i & wdwe == "Weekend")
        averagesteps2 <- mean(ds.int2$steps)
        averagesteps.list2 <- c(averagesteps.list2, averagesteps2) 
}
```

Then, we are going to combine the two results.

```{r}
steps <- c(averagesteps.list1, averagesteps.list2)
```

We also need to create new "interval" and "wdwe" columns. 

```{r}
interval <- c(ds$interval[1:288], ds$interval[1:288])
wdwe <- c(rep("weekday", 288), rep("weekend", 288))
```

Now we have all the components, we can create a new data.

```{r}
new.ds <- data.frame(interval = interval, steps = steps, wdwe = wdwe)
```

The new data looks something like this:
```{r}
head(new.ds)
tail(new.ds)
```

Finally, we can construct the **multi-panel time-series plot** using the lattice plotting system. 

```{r}
library(lattice)
xyplot(steps ~ interval | wdwe, data = new.ds, layout = c(1,2), type = "l", xlab = "Interval", ylab = "Number of steps")
```

Thank you for assessing my assignment! 
